Concurrent Programming with CXL

### 2020-06-24 book/paper.tex
Working on ABP chapter

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account.
In this chapter, we will consider two actors, a sender and
a receiver, that communicate by message passing over a network
that may lose messages.

>>>>

A distributed system is a concurrent system in which a collection
of processes communicate by message passing, much the same as
in the actor model.
The most important difference between distributed and concurrent
systems is that the former takes \emph{failures} into account.
In this chapter, we will consider two actors, Alice and Bob.
Alice wants to send a sequence of application messages to Bob,
but the underlying network may lose messages.
The network does not re-order messages: when sending messages
$m_1$ and $m_2$ in that order, then if both messages are received,
$m_1$ is received before $m_2$.
Also, the network does not create messages out of nothing: if
message $m$ is received, then message $m$ was sent.

We need a network protocol that Alice and Bob can run.
In particular, it has to be the case that if Alice sends application
messages $m_1, ..., m_n$ in that order, then if Bob receives an
application message
$m$, then $m = m_i$ for some $i$ and Bob will already have received
application messages $m_1, ..., m_i$ (safety).
Also, if the network is fair, then if Alice sends application message
$m$, then eventually Bob should deliver $m$ (liveness).

We assume that there are two unreliable network channels: \texttt{s\_chan}
from Alice to Bob, and \texttt{r\_chan} from Bob to Alice.
Because of the network's properties, we can model each
channel as a variable that either contains a network message or nothing
(we use \texttt{()} in the model).

The \emph{Alternating Bit Protocol} is suitable for our purposes.
Alice and Bob each maintain a zero-initialized sequence number,
\texttt{s\_seq} and \texttt{r\_seq} resp.
Alice sends a network message to Bob containing an
application message as \emph{payload} and Alice's sequence number
as \emph{header}.
When Bob receives such a network message, Bob returns an
an \emph{acknowledgment} to Alice, which is a network message
containing the sequence number in the message that Bob received.

In the protocol, Alice keeps sending the same network message until she
receives an acknowledgment with the same sequence number.
When this happens, Alice increments her sequence number.  She is now
ready to send the next message she wants to send to Bob.
Bob, on the other hand, waits until he receives a message matching
Bob's sequence number.  If so, Bob \emph{delivers} the payload in the
message and increments his sequence number.
Because of the network properties,  a one-bit sequence number suffices.
